---
title: "Analyzing-image-metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing-image-metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style>
p {
  text-align: justify;
}
</style>
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide'}
library(CSM)
```

## Index

- [Introduction](#introduction)
- [Test dataset info](#test-datasets)
- [Examples](#examples)
    - [Continuous metadata](#continuous-metadata)
    - [Qualitative metadata](#qualitative-metadata)
    - [Time-to-event analysis](#time-to-event-analysis)

## Introduction
CSM implements a master function to explore association of tissue features with tissue metadata. The function `Clinical_Data_analyzer()` allows the user to perform
association studies. It only requires two data sources. First, tissue metadata that has been adequately formatted using the `Clinical_Data_arrange_function()`.
Second, a summary tibble of the image features (for example cell densities by sample, heterogeneity by sample, neighborhood counts by sample...). Image features computed using
CSM are always numeric variables (instance counts, cell densities, spatial interaction metrics....). As far as each sample
is represented uniquely in both data sources (no duplicated entries in either data source are allowed), the function will be executed.

## Test datasets
For the following examples, we will use a dataset of endometrial carcinoma samples including 500K cells
obtained from 49 TMA samples. The cells have been assigned a phenotype label. The image metadata associated with this dataset
was randomly generated and includes information regarding patient age, survival status and tumor MMRP status.

```{r data, eval = TRUE}
CSM_PhenotypeTMA_test
CSM_ClinicalTMA_test
```

## Examples
For the following examples, we will first quantify cell densities by sample. 

```{r Cell_density, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
#Compute tissue area
DATA_AREA <-
 Image_size_calculator(
    DATA = CSM_PhenotypeTMA_test,
    Strategy = "Concave_hull",
    Hull_ratio = 0.4
)
#Obtain cell densities
Cells_by_sample <-
 Phenotype_quantifier(
    CSM_PhenotypeTMA_test,
    Calculate_Density = TRUE,
    DATA_Area = DATA_AREA
)

#Print the result
Cells_by_sample
```

In addition we need to arrange our image metadata to an adequate format using the `Clinical_Data_arrange_function()`.
```{r Clin_arrange, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
#Arrange clinical data
DATA_CLINICAL <-
  Clinical_Data_arrange_function(
     DATA = CSM_ClinicalTMA_test,
     Subject_Names = "Sample",
     Outcomes_to_keep = c("AGE", "MMRP_status", "DEATH", "OS_m")
)

#Print the result
DATA_CLINICAL
```


### Continuous metadata
Our first example will analyze the relationship between density of CD8 and macrophage cells and the patient's age. The function will return Pearson correlation
estimates and p values, as well as summary plots.
```{r Clin_continuous, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
Clinical_Data_analyzer(
   DATA = Cells_by_sample,
   DATA_var = c("Density_CD8", "Density_MACROPHAGE"),
   DATA_Clinical = DATA_CLINICAL,
   Clinical_var = "AGE",
   Perform_time_to_event = FALSE
)
```

### Qualitative metadata
Additionally, the function can work with qualitative data. Importantly, if qualitative data is encoded using numeric values, 
they must be turned to factor or character vectors before the analysis.
Otherwise, the function will analyze these variables as numeric and perform correlation analysis. 
The function will compute averages and dispersion metrics by group.
In addition t-tests or ANOVA will be computed. The summary graphs will include the non-parametric association test.


```{r Clin_qualitative, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
Clinical_Data_analyzer(
   DATA = Cells_by_sample,
   DATA_var = c("Density_CD8", "Density_MACROPHAGE"),
   DATA_Clinical = DATA_CLINICAL,
   Clinical_var = "MMRP_status",
   Perform_time_to_event = FALSE
)
```

### Time-to-event analysis
In order to perform time to event analysis, the image metadata must contain information of follow-up (a numeric vector) and event (also a numeric vector). 
The function will compute the optimal cut-off point based on pvalue optimization (using the `survminer::surv_cutpoint()` function). Once calculated, 
the function will fit a cox PH model and draw KM plots. The p value obtained using the log-rank test will be displayed in the KM plots.

```{r Clin_surv, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
Clinical_Data_analyzer(
   DATA = Cells_by_sample,
   DATA_var = c("Density_CD8", "Density_MACROPHAGE"),
   DATA_Clinical = DATA_CLINICAL,
   Perform_time_to_event = TRUE,
   Time_variable = "OS_m",
   Event_variable = "DEATH"
)
```


```{r Session_info, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
sessionInfo()
```
