---
title: "Cell phenotype labelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cell-phenotype-labelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
p {
  text-align: justify;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results= 'hide'}
library(CSM)
```

## Index

- [Introduction](#introduction)
- [Test dataset info](#test-datasets)
- [Phenotyping by positive feature pattern](#normalizing-using-mxnorm)
    - [Thresholding features](#feature-thresholding)
    - [Finding recurrent patterns](#finding-recurrent-patterns)
    - [Assign cell labels](#assign-cell-labels)
- [Unsupervised clustering](#unsupervised-clustering)
    - [Two-step clustering](#two-step-clustering)
- [CELESTA semi-supervised clustering](#celesta)
- [After cell phenotype labelling](#after-cell-phenotype-labelling)
    - [Phenotype evaluation app](#phenotype-evaluation-app)
    - [Concordance evaluation](#concordance-evaluation)
    - [Labelling by consensus](#labelling-by-consensus)
    - [Quantification & visualization](#quantification-&-visualization)

***
