---
title: "Cell phenotype labelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cell-phenotype-labelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
p {
  text-align: justify;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results= 'hide'}
library(CSM)
```

## Index

- [Introduction](#introduction)
- [Test dataset info](#test-datasets)
- [Phenotyping by positive feature pattern](#phenotyping-by-positive-feature-pattern)
    - [Thresholding features](#feature-thresholding)
    - [Finding recurrent patterns and assigning cell labels](#finding-recurrent-patterns-and-assigning-cell-labels)
- [Unsupervised clustering](#unsupervised-clustering)
    - [Two-step clustering](#two-step-clustering)
- [CELESTA semi-supervised clustering](#celesta)
- [Image-based supervised labeling](#image-based-supervised-labeling)
- [After cell phenotype labelling](#after-cell-phenotype-labelling)
    - [Phenotype evaluation app](#phenotype-evaluation-app)
    - [Concordance evaluation](#concordance-evaluation)
    - [Labeling by consensus](#labeling-by-consensus)
    - [Quantification & visualization](#quantification-and-visualization)

***

## Introduction
Cell phenotype labelling is a crucial step in spatial biology analysis. This process assigns
an identity label to every cell of a dataset. Many of the analysis pipelines in CSM and elsewhere
require cells to have an assigned phenotype label. CSM provides several ways to achieve labelling.
The choice between approaches is based on the characteristics of the cell feature dataset. If you cannot
decide between method the user can always perform [Labeling by consensus](#labeling-by-consensus).

* [Phenotyping by positive feature pattern](#phenotyping-by-positive-feature-pattern): Based on feature
  thresholding. Well suited for datasets with few features and target cell phenotypes.
* [Unsupervised clustering](#unsupervised-clustering): Based on various clustering algorithms. Well suited
  for datasets with many features. May miss infrequent cell phenotypes if parameters are not adequately tuned.
* [CELESTA semi-supervised clustering](#celesta): Based on CELESTA algorithm. Well suited if target cell phenotypes
  are known and user has prior information regarding expected feature expression of these cell phenotypes.
* [Image-based supervised labeling](#image-based-supervised-labeling): Based on manual labeling of cells based on
  images. Manually labeling cells is time consuming but adequate to identify cell phenotypes even in noisy scenarios.

## Test datasets
The test dataset includes cell feature matrix obtained from two 500x500 breast carcinoma images. 
They have been stained with a multiplexed immunofluorescence technique for the following markers: 
"DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3". 
The original images can be found at: https://immunoatlas.org/.

For the current example we will use information from around 1K cells. The mean non-normalized intensity of CK, CD8a and GZMB will be used.

```{r data, eval = TRUE}
CSM_Arrangedcellfeaturedata_test
```

## Phenotyping by positive feature pattern
This is one of the most popular methods to assign cell phenotypes. It is simliar to
marker expression gating in the field of flow cytometry. It requires setting thresholds for every feature.
Cells showing a feature intensity above the threshold will be considered positive for that specific feature. After 
thresholding feature positivity pattern can be used to assign a phenotype label to each cell. 

Feature thresholding is well suited for experiments with few measured features and
when the number cell phenotypes being identified is small. As the number of features
increases the resulting number of different positivity patterns can be prohibitively
large to work with. Users must understand that thresholding is usually a heavily supervised
approach. If a large number of features need to be included in the analysis, unsupervised, CELESTA or 
image-based supervised methods are preferred.

### Feature thresholding
Thresholding a feature basically means finding an adequate threshold to define positivity.
CSM implements a wide repertoire of automatic threshold identification algorithms. Most of them
are based on histogram analysis and try to reduce the within group variability compare to the
between group variability. Most of the algorithms included in CSM come derive from greyscale image 
binarization pipelines. 

Some of the approaches are very restrictive like RenyiEntropy or MaxEntropy while others are very 
tolerant like TriClass_Otsu or Huang. In addition, CSM implements methods that are not based on
histogram analysis (like mean, quantile or arbitrary thresholds).

In addition, thresholding can be executed in a LOCAL or in a GLOBAL manner. Local thresholding
will calculate a unique threshold for every image in a dataset. In contrast, global thresholding
will find a single threshold for all images in a dataset. In general, local thresholding is discouraged.
Local thresholding can potentially find thresholds according to image-specific noisy features. Although
global thresholding is also influenced by noise, it's more robust than local thresholding.

In order to explore thresholding approaches the user can use the `Thresholding_tester_app()`.
This app, can be used to explore images, cell ovearlays with feature intensity values and thresholding
approach results.

```{r app, eval = FALSE}
#Can only run interactively

Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#Deploy app----------------------------------------------
Thresholding_tester_app(
   DATA = CSM_Arrangedcellfeaturedata_test,
   Directory = Input_Dir,
   Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3")
)

#Remove directories---------------------------------------------------------
unlink(Input_Dir, recursive = TRUE)
```

Once a thresholding approach has been found to be appropriate, it's time to perform
thresholding. 

CSM deploys to feature thresholding methods. The first applies the same thresholding
strategy to all the features in the dataset and is executed using the `Thresholding_function()`.

```{r simple_thres, eval = TRUE}
DATA_Thresholded <- 
  Thresholding_function(
    DATA = CSM_Arrangedcellfeaturedata_test,
    Strategy = "EBI_Otsu",
    Local_thresholding = FALSE,
    Method_autothreshold = NULL,
    number_iterations_TriClass = NULL,
    Percentile = NULL,
    Defined_threshold = NULL,
    Levels = NULL
)
DATA_Thresholded
```

However, it is commonplace that different features require different thresholding approaches.
To address this situation, CSM implements `Thresholding_function_tailored()`. This function allows
the user to apply specific thresholding methods to every feature in a dataset. To this end,
the user must first generate a tibble (or dataframe) containing thresholding parameters for every feature being
processed:
```{r tailored_thres, eval = TRUE}
Thresholding_strategy_tibble <-
  tibble::tibble(
    variable = names(CSM_Arrangedcellfeaturedata_test)[-(1:4)],
    Strategy = c("TriClass_Otsu", "Kmeans", "Autothreshold"),
    Local_thresholding = F,
    Method_autothreshold = "Otsu",
    number_iterations_TriClass = 20,
    Percentile = NA,
    Defined_threshold = NA,
    Levels = NA
)

Thresholding_function_tailored(
  DATA = CSM_Arrangedcellfeaturedata_test,
  Variables_tibble = Thresholding_strategy_tibble
)
```

### Finding recurrent patterns and assigning cell labels
Once features have been thresholded, the next step is to find recurrent 
positivity patterns. These patterns are essentially the different cell phenotypes that
can be found in the dataset. CSM finds these patterns using the `Marker_combinator_generator()`.
The function can take a while to execute in very large datasets. The function will return
the possibilities arranged by the abundance in the dataset.

```{r Combin, eval = TRUE}
Phenotype_possibilities <- Marker_combinator_generator(
  DATA = DATA_Thresholded,
  Markers = names(DATA_Thresholded)[-c(1:4)]
)
Phenotype_possibilities
```

Now that all positivity combination patterns have been identified, the user must assign
a cell phenotype label to each pattern.

```{r Possibilities, eval = TRUE}
Phenotype_possibilities$Phenotype <-
  c('TUMOR', 'OTHER', 'CD8_GZMBneg', 'CD8_GZMBneg', 'OTHER', 'CD8_GZMBpos', 'CD8_GZMBpos')
Phenotype_possibilities[4:7]
```

Now we can combine this phenotype label information with the thresholded cell feature data
to assign the final cell phenotype labels.
```{r Phenotypes, eval = TRUE}
DATA_Phenotypes <-
  Phenotype_assigner_function(
    DATA = DATA_Thresholded,
    Phenotype_possibilities = Phenotype_possibilities
    )
table(DATA_Phenotypes$Phenotype)
```

## Unsupervised clustering
Unsupervised clustering is also a very useful method to perform cell phenotype labeling.
It can be specially useful in datasets with a high number of features. However, these methods
can be computationally intensive. To alleviate the computational burden, dimension reduction
and dataset denoising can be performed before clustering.

Here's a short description of clustering methods implemented in CSM:

* __Consensus Clustering and K_Means_Meta_clustering__: Well poised to find robust phenotypes. 
  It suffers with increased number of cells in the dataset (indeed with as few as 50K cells it can require huge amounts of RAM).
  K-means Meta clustering partially solves this issue. It first performs K-means to find a reduced number of centroids. These
  centroids can then be clustered using consensus clustering.
* __Self-organizing Maps (SOM)__: One of the most computationally efficient algorithms. It can usually miss low abundance cell types. It can also suffer
  with very small dataset (with les than 1K cells).
* __Graph based clustering__: Includes the very popular Louvain and Leiden algorithms. Computing graphs can take plenty of time and the results
  of the algorithm widely depend on the number of neighbors to build the KNN graphs and the resolution of the clustering method. However, these
  methods yield awesome results on datasets that have undergone dimension reduction
* __Batched K-means, Gaussian Mixture Models (GMM) and Clustering Large Applications (CLARA)__: These are the methods developed to deal with really large
  datasets. They perform reasonably well in both reduced and non-reduced scenarios. Stimating the right amount of cell types to identify can take some time.

Denoising can help to remove cells that have absent feature expression or that show intermediate phenotypes that
are hard to cluster. Denoising methods implemented in CSM are the following:

* __Quantile__: Removes cells that show feature expression below a desired quantile for ALL the features in the dataset.
* __Standard deviation__: Removes cells that show feature expression below the number of standard deviations from the mean for ALL the features
  in the dataset.
* __Threshold__: Removes cell that show no feature expression above the desired threshold for ALL the features in the dataset.
* __Otsu__: Performs Otsu thresholding for all the features. Cells without positivity in any of the features are removed from the analysis.
* __Density-based clustering on dimension reduction data__: Can only be performed if dimension reduction has been performed.
  Performs DBSCAN clustering. Unclustered cells (demonstrating undeterminate phenotypes) are removed from the analysis.

Once the clustering method, denosing and dimension reduction strategy have been selected clustering is straightforward.
Besides SOM and Graph-based clustering, all methods will try to estimate the adequate number of clusters to be computed.

```{r clustering, eval = TRUE, fig.align='center', out.width='100%', fig.width=6, fig.height=4}
Clustering_results <-
  Clustering_Phenotyper(
    DATA = CSM_Arrangedcellfeaturedata_test,
    Strategy = "Graph_Based",
    
    Apply_Denoise = TRUE,
    Denoising = "Otsu",
    
    Perform_Dimension_reduction = TRUE,
    Dimension_reduction = "UMAP",
    Dimension_reduction_prop = 1,
    Cluster_on_Reduced = TRUE,
  
    Nearest_neighbors_for_graph = 10,
    Graph_Method = "Leiden",
    Graph_Resolution = 0.05
)
```

After performing the clustering, according to the output graphs, the user can change the names of the clusters using the
`DATA_Phenotype_renamer()` function.

```{r re-namer, eval = FALSE}
#Can only run interactively

Clustering_results_DATA <-
  DATA_Phenotype_renamer(Clustering_results$DATA, 
                         c("OTHER", "TUMOR", "TUMOR", "CD8"))
table(Clustering_results_DATA$Phenotype)
```

### Two-step clustering
Clustering can be applied to already defined cell types based on previously unused features. Although re-clustering can be
applied to any phenotype labeling approach, it is reasonable to explain it here. This function reliefs the stress of the initial
clustering approach. The user can first assign core cell phenotype labels and then refine this initial core labels using the
`ReClustering_function()`. 

The re-clustering function works with a single cell phenotype at a time. It can re-cluster based on various features. 
It supports clustering methods like SOM, Batch K-means and GMM and thresholding methods like Arbitrary and Multilevel. Although not
formally evaluated by the function, features included in the re-clustering process should not have been used to define the initial core phenotype 
labels.

```{r Re-clustering, eval = FALSE}
#Can only run interactively

#Initial clustering based on CK and CD8a
Clustering_results <-
  Clustering_Phenotyper(
    DATA = CSM_Arrangedcellfeaturedata_test[-7],
    Strategy = "Graph_Based",
    
    Apply_Denoise = TRUE,
    Denoising = "Otsu",
    
    Perform_Dimension_reduction = TRUE,
    Dimension_reduction = "UMAP",
    Dimension_reduction_prop = 1,
    Cluster_on_Reduced = TRUE,
  
    Nearest_neighbors_for_graph = 10,
    Graph_Method = "Leiden",
    Graph_Resolution = 0.001
)

Clustering_results_DATA <-
  DATA_Phenotype_renamer(Clustering_results$DATA, 
                         c("OTHER", "TUMOR", "CD8"))

ReClustering_function(
    DATA = Clustering_results_DATA,
    DATA_aside = CSM_Arrangedcellfeaturedata_test,
    Phenotype_variable = "Phenotype",
    Phenotype_to_recluster = "CD8",
    Variables_to_recluster = "GZMB_AVERAGE",
    Strategy = "Multilevel",
    Levels = 3
    )

```

After re-clustering execution, the new labels can be modified at will using the `DATA_Phenotype_renamer()` function.

## CELESTA
The CELl typE identification with SpaTiAl information (CELESTA) algorithm was developed by the
Plevritis group (https://www.nature.com/articles/s41592-022-01498-z). In short, it assigns 
cell labels based on target cell phenotype information provided by the user. The algorithm runs a 
Gaussian Mixture Model to assign labels based on the information provided in sequential labeling rounds.

The algorithm is computationally efficient. However, in our experience, parameter tuning is key
to identify infrequent cell types. 

There are two basic steps to execute CELESTA in CSM. First, the user needs to generate a CELESTA template.
This is essentially a .csv file that needs to be fullfilled by the user with the target phenotypes information.
Each row represents a target phenotype. Columns provide expected feature expression patterns of target phenotypes
as well as parameters related to algorithm sensitivity and specificity. Information about completing the template
can be found at https://github.com/plevritis-lab/CELESTA. 

```{r template, eval = FALSE}
#Can only run interactively

CELESTA_template_generator(
   DATA = CSM_Arrangedcellfeaturedata_test,
   Markers_to_keep = c("CK_EPCAM_AVERAGE", "CD8a_AVERAGE", "GZMB_AVERAGE"),
   Template_name = "Mypath/name_of_the_template" #.csv at the end is not required, the function will add it for you
   )
```

After completing the template the CELESTA is executed using the `CELESTA_phenotyper()` function.
```{r CELESTA, eval = FALSE}
#Can only run interactively

DATA_Phenotypes <- 
  CELESTA_phenotyper(
    DATA = CSM_Arrangedcellfeaturedata_test, 
    Template_path = "Completed_template_path.csv",
    Alternative_CSV_locale = TRUE, 
    N_cores = 1, 
    Apply_filters = TRUE, 
    high_marker_threshold = 0.90, 
    low_marker_threshold = 0.10, 
    max_iteration = 10, 
    cell_change_threshold = 0.01
    )
```

Do not despair if results are not as expected. Obtaining satisfactory results using CELESTA is a matter
of trial and error. Try modifying anchor parameters and labeling round strategy until success.

## Image-based supervised labeling
This approach relies on the ability of the user to inspect images and manually label different cell types.
These labelled cells will be used to train a machine learning or deep learning algorithm. After training,
the resulting model can be used to label all cells in the dataset.

This approach is very powerful. It allows users to identify cells based on image findings. It can easily be trained
to identify low abundance or complex phenotypes. In addition, it allows the algorithm to use not only cell information
but also information of cell neighbors. This can be useful when cytoplasm segmentation is defective. However, it must be
considered that manual labeling of cells is time consuming, and the model performance may be highly dependent on the number
and quality of the cells selected for training.

In order to generate the model CSM implements a dedicated shiny App (`Image_based_phenotyper_App_launcher()`). The app will
automatically match the names of the image files and the Subject_Names in data. If missmatches are encountered a warning message
will pop-up before launch.

```{r model, eval = FALSE}
#Can only run interactively

#Create temporary directory
Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#Deploy app
Image_based_phenotyper_App_launcher(
   DATA = CSM_Arrangedcellfeaturedata_test,
   Directory = Input_Dir,
   Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
   Channels_to_keep = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3")
)

#Remove directories
unlink(Input_Dir, recursive = TRUE)
```

Using the App the phenotype label assign model can be created and saved in the Global Environment.
This model can then be used to phenotype all cells in the dataset using the `Model_cell_phenotyper()`.

```{r model_eval, eval = FALSE}
#Can only run interactively

print(CSM_ModelParam_test)

#Use the model to assign cell labels to all cells in the dataset
Model_cell_phenotyper(
 DATA = CSM_Arrangedcellfeaturedata_test,
 Model_parameters = CSM_ModelParam_test,
 N_cores = 1
)
```

## After cell phenotype labelling
Once cell phenotype labels have been assigned, users are encouraged to check that results are
satisfactory before proceeding. Flawed cell phenotype labels usually result in low quality
analyses. CSM implements tools to visually [evaluate](#phenotype-evaluation-app) phenotype labeling results, analyze
[concordance](#concordance-evaluation) between methods, reach a final label choice between methods by [consensus](#labeling-by-consensus)
and [quantifying](#quantification-and-visualization) results.

### Phenotype evaluation App
CSM implements a shiny App designed to evaluate phenotyping results. The app displays images, cell overlay with phenotype labels.
In addition, sample summaries are shown and user can evaluate the feature expression values for selected cells.

```{r Phenotype_app, eval = FALSE}
#Can only run interactively

Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#Deploy app----------------------------------------------
Phenotyping_evaluator_shiny_app_launcher(
    DATA = CSM_Phenotypecell_test,
    Directory = Input_Dir,
    Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3")
)

#Remove directories---------------------------------------------------------
unlink(Input_Dir, recursive = TRUE)
```

### Concordance evaluation
If concordance between phenotyping methods is high, which method is finally chosen is irrelevant.
In contrast, low concordance between methods may warrant further exploring the source of discrepancies.
To analyze concordance, CSM implements to related functions. First `Concordance_calculator()` can be used to compute
Rand or Fowlkes–Mallows concordance indexes between as many methods as desired. The function will return the median
p25 and p75 between sample concordance as well as a heatmap to spot samples with high or low concordance values.
Concordance evaluation will be fair as far as the cell phenotype labels are shared between the methods being compared.

We will illustrate this by randomly assigning cell phenotype labels and checking that the FM index is low given that there
are only 4 possible cell phenotypes in the dataset.

```{r Concordance, eval = TRUE, fig.align='center', out.width='100%', fig.width=12, fig.height=7}
library(magrittr)
library(dplyr)
#Generate datasets with random cell phenotype labels----------------------
CSM_Phenotypecell_test_1 <-
 CSM_Phenotypecell_test %>%
     mutate(Phenotype = sample(unique(CSM_Phenotypecell_test$Phenotype),
                               size = nrow(CSM_Phenotypecell_test),
                               replace = TRUE))
CSM_Phenotypecell_test_2 <-
 CSM_Phenotypecell_test %>%
     mutate(Phenotype = sample(unique(CSM_Phenotypecell_test$Phenotype),
                               size = nrow(CSM_Phenotypecell_test),
                               replace = TRUE))
CSM_Phenotypecell_test_3 <-
 CSM_Phenotypecell_test %>%
     mutate(Phenotype = sample(unique(CSM_Phenotypecell_test$Phenotype),
                               size = nrow(CSM_Phenotypecell_test),
                               replace = TRUE))

#Find concordance between results--------------------------------------
Concordance_calculator(
  Random_1 = CSM_Phenotypecell_test_1,
  Random_2 = CSM_Phenotypecell_test_2,
  Random_3 = CSM_Phenotypecell_test_3,
  Variable = "Phenotype",
  Strategy = "FM"
)
```


In addition, CSM implements the `Confusion_matrix_plotter()` function, that automatically draws a confusion matrix between
two different methods.
```{r Confusion, eval = TRUE, fig.align='center', out.width='100%', fig.width=12, fig.height=7}
Confusion_matrix_plotter(
  Random_1 = CSM_Phenotypecell_test_1,
  Random_2 = CSM_Phenotypecell_test_2,
  Variable = "Phenotype"
)
```

### Labeling by consensus
Although voting can be a flawed way of solving discrepancies (is better to analysze sources of conflict),it
can be a good way to solve minor discordance between methods. It can also be a good way to address method 
weaknesses by generating consensual choices. CSM implements consensus labeling through the 
`Consensus_phenotype_assigner()` function. The function supports different ways of analyzing votes, as well as
providing specific weights to highly confidence methods.

```{r consensus, eval = FALSE, fig.align='center', out.width='100%', fig.width=12, fig.height=7}
#Can only run interactively
DATA_Phenotypes_consensus <-
  Consensus_phenotype_assigner(
    Random_1 = CSM_Phenotypecell_test_1,
    Random_2 = CSM_Phenotypecell_test_2,
    Random_3 = CSM_Phenotypecell_test_3,
    
    Win_strategy = "Majority",
    Weights = c(2,1,1),
    No_consensus_value = NA_character_,
    Tie_break_method = "Random",
    N_cores = 1
 )
```

### Quantification and visualization
Once the phenotype labeling process is considered to be adequate, it's a good idea to quantify cell phenotypes.
This can be easily done with the `Phenotype_quantifier()` function.

```{r quantify, eval = TRUE, fig.align='center', out.width='100%', fig.width=12, fig.height=7}
#Calculate tissue size 
DATA_AREA <-
  Image_size_calculator(
    DATA = CSM_Phenotypecell_test,
    Strategy = "Tiling",
    Tile_accuracy = 100
    )

#Calculate the cell counts by image-------
Phenotypes_by_Sample <-
  Phenotype_quantifier(
    DATA = CSM_Phenotypecell_test,
    Calculate_Density = TRUE,
    DATA_Area = DATA_AREA
)
Phenotypes_by_Sample
```

In addition, cell composition for every image can be used to obtain a barplot with the relative percentages of each cell type.
```{r barplot, eval = TRUE, fig.align='center', out.width='100%', fig.width=12, fig.height=7}
Barplot_generator(
    DATA = Phenotypes_by_Sample,
    Phenotypes_included = c("TUMOR", "CD8_GZMBpos", "CD8_GZMBneg"),
    Ordering_phenotype = "CD8_GZMBpos"
)
```

Moreover, cells from individual images can be plotted with or without plotting the image.
```{r images, eval = TRUE, fig.align='center', out.width='100%', fig.width=12, fig.height=10}
#Without images
Cell_plotter(
   DATA = CSM_Phenotypecell_test,
   Image_name = "ABCM22001_B14_MiniCrop.tif",
   Phenotypes_included = c("TUMOR", "CD8_GZMBpos", "CD8_GZMBneg", "OTHER")
)

#With images
Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#Draw the plot
Cell_image_plot_generator(
    Image_directory = dir(Input_Dir, full.names = TRUE)[2],
    Channel_to_display = 5,
    Image_rotate = NULL,
    Image_x_flip = FALSE,
    Image_y_flip = TRUE,
    Gamma_level = 0,
    Equalize = FALSE,
    Black_level = 1,
    White_level = 20,

    DATA = CSM_Phenotypecell_test,
    Image_name ="ABCM22001_B14_MiniCrop.tif",
    Color_by = "Phenotype",
    Point_size = 2,
    Pixel_distance_ratio = NULL
)

#Remove directories---------------------------------------------------------
unlink(Input_Dir, recursive = TRUE)
```

```{r Session_info, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
sessionInfo()
```
