---
title: "Image-QC-flagging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Image-QC-flagging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
p {
  text-align: justify;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide'}
library(CSM)
```

## Index
- [Introduction](#introduction)
- [Test dataset info](#test-datasets)
- [Background and examples](#background-and-examples)

## Introduction

Checking if images or image-channels have any quality issues is an essential, yet unresolved aspect of image analysis. 
Signal that is not derived from true presence of a target tissue bio-molecule is commonly referred as 'noise'. Noise is commonplace
in spatial proteomics and transcriptomics. 

## Test datasets

The test datasets consists of two small 500x500 pixels breast carcinoma TIFF images. They have been 
stained with a multiplexed immunofluorescence protocol for the following markers: 
"DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3". 
The original images can be found at: https://immunoatlas.org/.


## Background and examples

The sources of noise are variable:

* __Pre-analysis:__ Include tissue processing issues (cold ischemia, fixation, freezing and thawing...), tissue folding artifacts or presence of
noise-producing cell types (for example erythrocytes show high peroxidase activity or cells rich in cytochromes show high autofluorescence).
* __Analysis:__ Include issues with antibody-antigen non-specific binding (for antibody-based protein detection methods), probe non-specific binding (for hibridization-based methods),
Or issues with analytical contaminants (like heavy metals in mass cytometry based analyses)
* __Imaging:__ Includes issues with illumination (non-even illumination), under or over-exposure, focusing issues, issues with light filters, image sensor overheating or image stiching artifacts.

Perfect images contain only true biological signal and no noise. However all noisy images are biased __in their own way__. This means that there is no universal way of quantifying and correcting
image noise. 

In stead of taking into account all potential sources of image noise (which would require several modelling approaches), CSM identifies gross quality issues based on the 
pixel information contained in the images. In every biological system there is a degree of expected spatial correlation (co-localiation) of biomolecules. For example, if two proteins
have a nuclear subcellular localization, they are expected to be spatially correlated. 

The `Channel_Sample_diagnostics()` function implemented in CSM is founded on the prior knowledge of biomolecule co-localization. In order to generate adequate results the user must provide
a list of expected correlation hubs (image channels containing protein information that are expected to show spatial correlation).

```{r Correlation hubs, eval = FALSE}
#Generate list of correlation hubs
Correlation_hubs <- list(Nuclear_markers = c("DAPI", "FOXP3"),
                        Tumor_markers = c("CK-EPCAM", "PDL1"),
                        Lymphocyte_markers = c("CD8a", "GZMB", "PD1"))
```

Afterwards, spatial autocorrelation can be tested. `Channel_Sample_diagnostics()` works also with tiled images. In order to ameliorate computational burden, images can be downsized to reduce the 
amount of pixels. Downsizing also allows the identification of more gross quality issues, reducing the chances of overestimating minor non-significant noise. The function will first compute a tissue
mask in a similar way as `Pixel_Threshold_calculator()` does. Spatial correlation will be computed using foreground pixels.
Although the function has been developed to work with images, `Channel_Sample_diagnostics()` can also deal with cell feature matrices, identifying spatial correlation on cell data in stead of pixels.

```{r Evaluation, eval = FALSE}
#Create temporary input and output directories------------------------------
Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
          function(Image){
            EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
                               file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
          })

#Run the test
Channel_Sample_diagnostics(
  N_cores = 1,
 
  Directory = Input_Dir,
  Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
  Pixel_downsize = 0.2,
  Channel_for_test = "DAPI",
  Channels_to_keep = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
  Threshold_type_tissueMask = "Arbitrary",
  Threshold_value_tissueMask = 0.05,
  Blurr_tissueMask = FALSE,
  Sigma_tissueMask = NULL,
 
  DATA = NULL,
 
  List_correlation_hubs = Correlation_hubs,
  Spearman_threshold = 0.05
)

#Example with cell feature matrix

```

Given the small size of the dataset it is hard to identify robust quality issues. However, it is worth mentioning that quality issues raised by `Channel_Sample_diagnostics()` should not be given for
granted without proper user inspection of problematic images or channels.
