---
title: "Image-QC-flagging"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Image-QC-flagging}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
p {
  text-align: justify;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide'}
library(CSM)
```

## Index
- [Introduction](#introduction)
- [Test dataset info](#test-datasets)
- [Examples](#examples)

## Introduction
Checking if images or image-channels have any quality issues is an essential, yet unresolved aspect of image analysis. 
Signal that is not derived from true presence of a target tissue bio-molecule is commonly referred as 'noise'. Noise is commonplace
in spatial proteomics and transcriptomics. 

The sources of noise are variable:

* __Pre-analysis:__ Include tissue processing issues (cold ischemia, fixation, freezing and thawing...), tissue folding artifacts or presence of
noise-producing cell types (for example erythrocytes show high peroxidase activity or cells rich in cytochromes show high autofluorescence).
* __Analysis:__ Include issues with antibody-antigen non-specific binding (for antibody-based protein detection methods), probe non-specific binding (for hibridization-based methods),
Or issues with analytical contaminants (like heavy metals in mass cytometry based analyses)
* __Imaging:__ Includes issues with illumination (non-even illumination), under or over-exposure, focusing issues, issues with light filters or image sensor overheating.



However, identifying and quantifying 

  

