---
title: "Working with RGB images"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working-with-RGB-images}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide'}
library(CSM)
```

## Index

- [Introduction](#introducción)
- [Color channel extraction APP](#test-datasets)
- [Performing channel extraction](#check-computing-resources-available)
- [Cell segmentation based on channels](#chopping-large-images)

## Introduction
RGB color images are another source of image information. CSM provides tools to extract different channels from
RGB color images according to different colors. This can be used to separate hematoxylin and eosin channels from 
conventional HE stained tissues, or to obtain hematoxylin and Diaminobenzidine color channels from immunohistochemical images.
These channels can be used for subsequent pixel based analysis or even to perform cell segmentation if adequate nuclear
channels can be obtained.

## The channel extraction testing APP
CSM implements a dedicated shiny App to explore how color channels can be extracted and generate the parameters that can be then applied
to all the images in the dataset.

