---
title: "Working with large images and TMAs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working-with-large-images-and-TMAs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
p {
  text-align: justify;
}
</style>


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide'}
library(CSM)
```

## Index

- [Introduction](#introduction)
- [Test dataset info](#test-datasets)
- [Check computing resources](#check-computing-resources-available)
    - [Chopping large images](#chopping-large-images)
    - [Cell segmentation using tiles](#cell-segmentation-using-tiled-images)
- [Pixel thresholding using tiles](#pixel-thresholding-of-tiled-images-and-re-assembled-images)
- [Downsizing large images](#downsizing-large-images)
- [De arraying TMA whole slide images into individual core images](#de-arraying-tmas)

***

## Introduction
Working with large images can be painful in R. Whole slide images can exceedingly large (up to tens of Gigabytes). This
can easily saturate RAM memory. Indeed many popular image handling R packages (like EBImage or magick) have image size
limits that can be easily exceeded by whole slide images. 

CSM has inbuilt pipelines to deal with this situation. CSM implements functions to chop large images into manageable smaller crops. 
After image processing (either cell segmentation, cell phenotype labeling or image pixel thresholding), cell or image 
pixel information can be re-assembled to match the original whole slide coordinates.

In addition, CSM has in-built tools to downsize large images to more manageable smaller sized images. These downsized images can be used for
plotting purposes or to feed the `TMA_dearrayer_function()` (see below).


## Test datasets

The test datasets consists of two small 500x500 pixels breast carcinoma images. They have been stained with a multiplexed
immunofluorescence protocol for the following markers: "DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3". 
The original images can be found at: https://immunoatlas.org/.

This is a trivial test dataset, as it contains very small images that could be easily handled by CSM. However, the principles
described in this example apply to larger whole-slide images.

```{r Images, eval = FALSE, results='hide'}
#Can only run in an interactive setting
#Generate a temporary directory to place the input images
Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})
```


## Check computing resources available
It is always a good idea to first check the CPU and RAM available.

```{r check, eval = TRUE}
Describe_my_computing_unit()
```

## Chopping large images
`Image_tile_deconstruction_function()` divides all the images in a directory into tiles. Accepted file extension include TIFF and OME.TIFF.
User can select the tile size (in pixels),the image channels to keep and the tile overlap. The smaller the tile size the easier it will 
be to work with the resulting chopped images, at the cost of an increased total tile number. If cell segmentation is going to be performed, 
tile overlap may be applied to account for segmentation artifacts due to edge proximity. Cell redundancy will be corrected in subsequent steps.
If RGB images need to be chopped, all RGB channels will be kept. The image chopping process is powered by the RBioFormats bioconductor R package. 

```{r chop, eval = FALSE}
#Can only run in an interactive setting
Input_Dir <- tempfile(pattern = "tempdir1_Input")
Output_Dir <- tempfile(pattern = "tempdir2_Output")
dir.create(Input_Dir, recursive = TRUE)
dir.create(Output_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#Perform tiling---------------------------------------------------
Image_tile_deconstruction_function(
   Directory = Input_Dir,
   Output_directory = Output_Dir,
   Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
   Channels_to_keep = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
   RGB_Color_images = FALSE,
   Tile_pixel_size = 250,
   Tile_Overlap = 50,
   N_cores = 1
)

#Check that files have been created-----------------------------------------
list.files(Output_Dir)
```

## Cell segmentation using tiled images
Cell segmentation and feature extraction proceeds as usual.
```{r Segment, eval = FALSE}
#Can only run in an interactive setting
Cell_DATA_tiles <- 
  Cell_segmentator_quantificator(
    Directory = Output_Dir,
    Parameter_list = CSM_SegmentParams_test,
    N_cores = 1,
    quantiles_to_calculate = 0.5
)
```

After cell segmentation has been performed, the dataset needs to be formatted using the `Data_arrange_function()`. 
Then, original cell coordinates can be obtained by running `Tile_to_image_cell_arrange_function()`. If tile overlap is detected, 
redundant cells (sharing spatial coordinates) will be removed automatically. In addition, 
if there is tile overlap, cells close to a shared tile edge can be removed according to
user preferences (with the 'Dist_to_edge' argument). The function works based on the information contained in the file name.
Manipulating the name can result in unexpected results. __PLEASE AVOID FILE NAME MODIFICATIONS!__
```{r Reassign_coordinates, eval = FALSE}
#Can only run in an interactive setting
#Arrange data
Cell_DATA_tiles <- 
  Data_arrange_function(
    DATA = Cell_DATA_tiles,
    X = 'm.cx',
    Y = "m.cy",
    Subject_Names = "imageID",
    Markers_to_keep = c("CK-EPCAM_AVERAGE", "CD8a_AVERAGE", "GZMB_AVERAGE")
)

#Return to original image X Y coordinates
Cell_DATA_WholeImage <-
  Tile_to_image_cell_arrange_function(
    DATA = Cell_DATA_tiles,
    Dist_to_edge = 20
)

#Check that results are OK
Cell_DATA_WholeImage %>% 
  dplyr::filter(Subject_Names == "ABCM22001_B09_MiniCrop.tiff") %>%
  ggplot2::ggplot(ggplot2::aes(x = X, y = Y)) + ggplot2::geom_point()

```

## Pixel thresholding of tiled images and re-assembled images
Single channels can be pixel-thresholded using image chunks. Afterwards thresholded image chunks can be used to re-assemble a 
thresholded whole image using the `Image_from_tile_rebuilder()` function. Since image chunks will be used in the thresholding process,
LOCAL thresholding using automatic algorithms can be extremely problematic (high risk of setting thresholds according to local artifacts).
Therefore, GLOBAL thresholding is highly recommended.

If images are being chopped to be thresholded, overlap between images is not supported. Although no error will be thrown by the 
related functions, the re-build process of overlapping tiles will result in aberrant re-assembled images.
```{r Tile_threshold, eval = FALSE}
#Can only run in an interactive setting
Input_Dir <- tempfile(pattern = "tempdir1_Input")
Output_Dir_tiles <- tempfile(pattern = "tempdir2_Output")
Output_Dir_tiles_thresholded <- tempfile(pattern = "tempdir3_Output")
Output_Dir_reconstruction <- tempfile(pattern = "tempdir4_Output")
dir.create(Input_Dir, recursive = TRUE)
dir.create(Output_Dir_tiles, recursive = TRUE)
dir.create(Output_Dir_tiles_thresholded, recursive = TRUE)
dir.create(Output_Dir_reconstruction, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#Divide images into tiles------------------------------------------------------
Image_tile_deconstruction_function(
   Directory = Input_Dir,
   Output_directory = Output_Dir_tiles,
   Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
   Channels_to_keep = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
   RGB_Color_images = FALSE,
   Tile_pixel_size = 250,
   Tile_Overlap = 0,
   N_cores = 1
)

#Threshold the CK_EPCAM channel of tiles
Pixel_Threshold_calculator(
     N_cores = 1,
     Directory = Output_Dir_tiles,
     Ordered_Channels = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
     Channels_to_keep = c("DAPI", "PDL1", "GZMB", "PD1", "CK-EPCAM", "CD8a", "FOXP3"),
     Target_channel = "CK-EPCAM",

     Save_processed_images = TRUE,
     Output_Directory = Output_Dir_tiles_thresholded,

     Local_thresholding = FALSE,
     Threshold_type = "Arbitrary",
     Threshold_value = 0.02,

     Threshold_type_tissueMask = "Absolute",
     Threshold_value_tissueMask = 0.001,
     Blurr_tissueMask = TRUE,
     Sigma_tissueMask = 0.5,
     Blurr_target = FALSE
)

#Rebuild images from tiles-----------------------
Image_from_tile_rebuilder(
   Directory = Output_Dir_tiles_thresholded,
   Output_directory = Output_Dir_reconstruction,
   RGB_Color_images = FALSE,
   N_cores = 1
)

#Check the files created-------------------------------------------------
list.files(Output_Dir_reconstruction)

#Remove directories---------------------------------------------------------
unlink(c(Input_Dir, Output_Dir_tiles, Output_Dir_tiles_thresholded, Output_Dir_reconstruction), recursive = TRUE)
```

## Downsizing large images
The CSM function `Smart_image_importer()` is capable of importing to the R session a downsized version of very large images.
The user must first provide the path to the image and the desired maximum number of pixels (in log10 scale) that the output
image will have (1 million max pixels by default).

In addition, the user can subset the image by cropping it or select specific image channels to be kept. The output image can be saved by
setting the `Save_processed_images = TRUE`.

How does the `Smart_image_importer()` work? First, the function will obtain image metadata (file type, pixel size, number of channels...).
If the image is smaller than 1.5Gb in size, it will attempt importing the image directly followed by downsizing. If the image is larger
than 1.5Gb it will sequentially import image chunks(tiles) and downsized them. Afterwards it will merge the chunks(tiles) back into a single
image. The function is designed to generate tiles that are around 0.5Gb in size.

```{r Smart_importer, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
#Create temporary input directory------------------------------
Input_Dir <- tempfile(pattern = "tempdir1_Input")
dir.create(Input_Dir, recursive = TRUE)

#Save images in Input directory
purrr::map(1:2,
function(Image){
   EBImage::writeImage(CSM_MiniMultiTiff_test[[Image]],
   file.path(Input_Dir, names(CSM_MiniMultiTiff_test)[Image]))
})

#reduce the pixel content of the image
Image <- Smart_image_importer(Image_directory = dir(Input_Dir, full.names = TRUE)[1], 
                     Log10_pixel_output = 5
)

print(Image)
```

Downsized images can be subsequently used for plotting purposes (for example using the `Image_channel_color_merger()` to combine several
channels into a single RGB color image). In addition, the `Smart_image_importer()` is also the function used by several Shiny APPs
implemented in CSM to import large images.

## De-arraying TMAs
Tissue micro-arrays (TMAs) are usually large whole slide images that require individual core image extraction before proceeding with 
analyses. The `TMA_dearrayer_function()` serves this purpose. It requires a downsized version of the whole slide TMA image. The function will use 
this downsized image to identify where tissue cores are located.

* __Tissue mask generation:__ First, the image is thresholded to find foreground (tissue) and background pixels. Foreground pixels will be used in
subsequent processing. To this end, the user may choose different strategies (identical to the approaches for tissue mask generation implemented
in `Pixel_Threshold_calculator()`). In our experience, it is best to use the arbitrary method and try various cut-off values to select the most
appropriate. 

* __Tissue mask processing:__ After obtaining the binary (foreground/background) pixel mask, the image will be processed to fill wholes. Then, erosion
followed by dilation can be applied according to user preferences. These pre-processing steps polish the initial mask and ease out the subsequent
watershed algorithm that will be applied to identify individual core areas. The choices of erosion and dilation round number and kernel size will be 
guided by the amount of noise in the image, the presence of tissue detachment or crush artifact and the size of the cores.

* __Tissue mask watersheding:__ The processed tissue mask is the subjected to a watershed algorithm to identify individual tissue core areas. To this end
the mask is first transformed to a distance map were the value of foreground pixels is equal to the distance of every pixel to the closest background
pixel. This allows the watershed algorithm to separate cores even when there are touching edges (that may occur in heavily populated TMAs). Two 
parameters influence the results of watershed algorithms in the case there are touching TMA cores: tolerance and extension (corresponding to
the `Watershed_tolerance` and `Watershed_ext` function arguments). High tolerance values will 
result in merging of TMA cores. The same is true for the extension parameter. Keeping the tolerance low (below 1) and low extension values (below 10) 
maximize the number of cores identify, at the cost of sometimes splitting a single tissue core into several pieces. At the end of this process the
function will return the identified cores and the user will have the opportunity to stop the function and refine parameters if required.

* __Estimate the number of rows and columns:__ The function will then compute the X and Y centroid coordinates for every core and will determine the 
number of rows and columns present in the TMA and the row and column index where every core belongs. This process is relevant as it will be used to label
each core after extraction, and if the user provides a matrix of core names, the row/col information will be used to match cores and names provided.
To this end, the function will use hierarchical clustering to group the core coordinates. This process can be refined modifying the `Tolerance_column`
and `Tolerance_row` arguments. Small tolerance values require striking distances between individual rows and columns to be identified. After performing
this step, the function will return the bounding box and provisional labels of the identified cores and the user may have the opportunity to stop the 
function and refine parameters if required.

* __Add user provided core labels:__ This is an optional by __HIGHLY ENCOURAGED__ step. User can provide a matrix or a dataframe/tibble containing 
core labels. The function will take this information, match the matrix according to the number of rows/columns of the matrix and the identified cores, 
and proceed to label the cores according to user preferences. Importantly, the matrix must not contain duplicated names. Cores that are known by the 
user to be empty should be indicated using NA_value. After performing labelling, the function will return the results of the alignment. The user can then
switch rows or column order or transverse the provided matrix at will until the labels match exactly with the tissue cores.

* __Extract tissue core images:__ Finally, the function will compute the bounding box of the cores in the original image. Individual TMA cores will
be subsetted from the original image using the `Smart_image_importer()`. Each image will be saved in the desired output directory.

```{r TMA_dearray, eval = FALSE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
TMA_image_downsized <-
    Smart_image_importer(Image_directory = "Path/to/original/TMA/Image", 
                         Log10_pixel_output = 6
  ) 
  
TMA_dearrayer_function(Original_Image_Directory = "Path/to/original/TMA/Image", 
                      Downsized_Image = TMA_image_downsized, 
                      Output_Directory = "Path/to/ouput/directory")

```

```{r Session_info, eval = TRUE, fig.align='center', out.width='100%', fig.width = 10, fig.height = 8}
sessionInfo()
```
